Mechanisms of AIAN Inequality in Child Welfare Across U.S. States.

### 2020-12-16 cfp_final_report.Rmd
update with rmd changes

```{r read-make}
#### make data for easy load
# source("make_state_tables.R")
# source("make_national_tables.r")
# source("read_adopt.r")
# 
# save.image("./cfp_final_report.RData")

@@

### ICWA AAIA data STATES

icwa_fc<-icwa_fc %>% 
  filter(!is.na(fc))

@@

icwa_nat<-icwa %>% 
  filter(!(is.na(fc_aian))) %>% 
  summarize(fc_aian = sum(fc_aian),
            fc_non = sum(fc_non),
            pop_aian21 = sum(pop_aian21),
            pop_non = sum(pop_nonAIAN21))

icwa_st<-icwa %>% 
  filter(!(is.na(fc_aian)))

@@

```{r nat-inv-age, fig.cap = "Age specific risk of experiencing first CPS investigation by state, 2014 - 2018 risk levels"}
ggplot(investigation_tables %>% 
         group_by(state, race_ethn, age) %>% 
         summarise(qmax = max(q_c), qmin=min(q_c), q = mean(q_c)),
       aes(x = age, y = q, ymin = qmin, ymax = qmax,
           color = race_ethn, fill = race_ethn)) + 
  geom_line() + 
  geom_ribbon(alpha = 0.5, color = NA) + 
  facet_geo(~state) + 
  labs(color = "", fill = "") + 
  theme_minimal() + 
  labs(x = "Age", y = "Proportion of children") + 
  theme(strip.text.x = element_text(size = 7),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 4)) + 
  scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) + 
  scale_fill_brewer(palette = "Dark2")+
  scale_color_brewer(palette = "Dark2")

@@

```{r nat-fc-age, fig.cap = "Age specific probability of entering foster care, 2014 - 2018 risk levels"}
ggplot(fc_tables %>% 
         group_by(state, race_ethn, age) %>% 
         summarise(qmax = max(q_c), qmin=min(q_c), q = mean(q_c)),
       aes(x = age, y = q, ymin = qmin, ymax = qmax,
           color = race_ethn, fill = race_ethn)) + 
  geom_line() + 
  geom_ribbon(alpha = 0.5, color = NA) + 
  facet_geo(~state) + 
  labs(color = "", fill = "") + 
  theme_minimal() + 
  labs(x = "Age", y = "Proportion of children") + 
  theme(strip.text.x = element_text(size = 7),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 4)) + 
  scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) + 
  scale_fill_brewer(palette = "Dark2")+
  scale_color_brewer(palette = "Dark2")
```

@@

map_dat<-investigation_tables_c %>% 
  filter(race_ethn=="AIAN") %>% 
  group_by(state) %>% 
  summarize(c = mean(c_c))

@@

map_dat<-subst_tables_c %>% 
  filter(race_ethn=="AIAN") %>% 
  group_by(state) %>% 
  summarize(c = mean(c_c))

@@

map_dat<-fc_tables_c %>% 
  filter(race_ethn=="AIAN") %>% 
  group_by(state) %>% 
  summarize(c = mean(c_c))

@@

map_dat<-tpr_tables_c %>% 
  filter(race_ethn=="AIAN") %>% 
  group_by(state) %>% 
  summarize(c = mean(c_c))

@@

map_dat_wht<-map_dat %>% 
  ungroup() %>% 
  filter(race_ethn == "White") %>% 
  rename(c_white = c_c) %>% 
  select(-race_ethn, -c, -race_ethn)

map_dat<-map_dat %>% 
  filter(race_ethn=="AIAN") %>% 
  left_join(map_dat_wht) %>% 
  mutate(disp = c_c / c_white)

@@

mn<-investigation_tables %>% 
  filter(.imp==1, age == 0, state == "MN") %>% 
  select(var, race_ethn, pop_max, q_c) %>% 
  mutate(name = "Investigations") %>% 
  bind_rows(
  subst_tables %>% 
  filter(.imp==1, age == 0,  state == "MN") %>% 
  select(var, race_ethn, pop_max, q_c) %>% 
  mutate(name = "Confirmed maltreatment"))

#### make descriptive table
mn<-mn %>% 
  filter(race_ethn=="AIAN") %>% 
  mutate("Number of children" = round(var/5,0),
         "Child population" = round(pop_max/5,0),
         "Event rate" = round(`Number of children`/ `Child population`,2)) %>% 
  select(name, `Number of children`, `Child population`, `Event rate`)

@@

```{r sub-inv}
ncands_sub_inv_tab<-subst_tables %>%
  rename(sub = q_c) %>% 
  select(.imp , age, race_ethn, state, sub) %>% 
  left_join(investigation_tables %>% 
              rename(inv = q_c) %>% 
              select(.imp, age, race_ethn, state, inv)) %>% 
  mutate(sub_inv = sub / inv) %>% 
  filter(sub_inv<1,
         !(is.na(sub_inv)))

@@

```{r fc-inv}
fc_cond_inv_tabs<-fc_post_inv_tables %>%
  rename(fc_inv = q_c) %>% 
  select(.imp,age,  race_ethn, state, fc_inv) %>% 
  left_join(investigation_tables %>% 
              rename(inv = q_c) %>% 
              select(.imp, age, race_ethn, state, inv)) %>% 
  mutate(fc_cond_inv = fc_inv / inv)

@@

fc_cond_sub_tabs<-fc_sub_tables %>%
  rename(fc_sub = q_c) %>% 
  select(.imp,age,  race_ethn, state, fc_sub) %>% 
  left_join(subst_tables %>% 
              rename(sub = q_c) %>% 
              select(.imp, age, race_ethn, state, sub)) %>% 
  mutate(fc_cond_sub = fc_sub / sub)

@@

```{r non-icwa}
non_icwa_cond_fc<-non_icwa_tables %>%
  rename(non_icwa = q_c) %>% 
  select(.imp, age, race_ethn, state, non_icwa) %>% 
  left_join(fc_tables%>% 
              rename(fc = q_c) %>% 
              select(.imp, age, race_ethn, state, fc)) %>% 
  mutate(non_icwa_fc = non_icwa / fc)

>>>>

```{r read-make}
# #### make data for easy load
# source("make_state_tables.R")
# source("make_national_tables.r")
# source("read_adopt.r")
# 
# save.image("./cfp_final_report.RData")

@@

@@

icwa_nat<-icwa_fc %>%
  group_by(race_ethn) %>% 
  summarize(fc = sum(fc, na.rm=T),
            adopted = sum(adopted, na.rm=T),
            boarding = sum(boarding, na.rm=T))

@@

```{r nat-inv-age, fig.cap = "Age specific risk of experiencing first CPS investigation by state, 2014 - 2018 risk levels"}
ggplot(investigation_tables %>% 
         group_by(state, race_ethn, age) %>% 
         summarise(qmax = max(q), qmin=min(q), q = mean(q)),
       aes(x = age, y = q, ymin = qmin, ymax = qmax,
           color = race_ethn, fill = race_ethn)) + 
  geom_line() + 
  geom_ribbon(alpha = 0.5, color = NA) + 
  facet_geo(~state) + 
  labs(color = "", fill = "") + 
  theme_minimal() + 
  labs(x = "Age", y = "Proportion of children") + 
  theme(strip.text.x = element_text(size = 7),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 4)) + 
  scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) + 
  scale_fill_brewer(palette = "Dark2")+
  scale_color_brewer(palette = "Dark2")

@@

```{r nat-fc-age, fig.cap = "Age specific probability of entering foster care, 2014 - 2018 risk levels"}
ggplot(fc_tables %>% 
         group_by(state, race_ethn, age) %>% 
         summarise(qmax = max(q), qmin=min(q), q = mean(q)),
       aes(x = age, y = q, ymin = qmin, ymax = qmax,
           color = race_ethn, fill = race_ethn)) + 
  geom_line() + 
  geom_ribbon(alpha = 0.5, color = NA) + 
  facet_geo(~state) + 
  labs(color = "", fill = "") + 
  theme_minimal() + 
  labs(x = "Age", y = "Proportion of children") + 
  theme(strip.text.x = element_text(size = 7),
        axis.text.x = element_text(size = 6),
        axis.text.y = element_text(size = 4)) + 
  scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) + 
  scale_fill_brewer(palette = "Dark2")+
  scale_color_brewer(palette = "Dark2")
```

@@

map_dat<-investigation_tables_c %>% 
  filter(race_ethn=="AIAN") %>% 
  group_by(state) %>% 
  summarize(c = mean(c))

@@

map_dat<-subst_tables_c %>% 
  filter(race_ethn=="AIAN") %>% 
  group_by(state) %>% 
  summarize(c = mean(c))

@@

map_dat<-fc_tables_c %>% 
  filter(race_ethn=="AIAN") %>% 
  group_by(state) %>% 
  summarize(c = mean(c))

@@

map_dat<-tpr_tables_c %>% 
  filter(race_ethn=="AIAN") %>% 
  group_by(state) %>% 
  summarize(c = mean(c))

@@

map_dat_wht<-map_dat %>% 
  ungroup() %>% 
  filter(race_ethn == "White") %>% 
  rename(c_white = c) %>% 
  select(-race_ethn)

map_dat<-map_dat %>% 
  filter(race_ethn=="AIAN") %>% 
  left_join(map_dat_wht) %>% 
  mutate(disp = c / c_white)

@@

mn<-investigation_tables %>% 
  filter(.imp==1, age == 0, state == "MN") %>% 
  select(var, race_ethn, pop, q) %>% 
  mutate(name = "Investigations") %>% 
  bind_rows(
  subst_tables %>% 
  filter(.imp==1, age == 0,  state == "MN") %>% 
  select(var, race_ethn, pop, q) %>% 
  mutate(name = "Confirmed maltreatment"))

#### make descriptive table
mn<-mn %>% 
  filter(race_ethn=="AIAN") %>% 
  mutate("Number of children" = round(var/5,0),
         "Child population" = round(pop/5,0),
         "Event rate" = round(`Number of children`/ `Child population`,2)) %>% 
  select(name, `Number of children`, `Child population`, `Event rate`)

@@

```{r sub-inv}
ncands_sub_inv_tab<-subst_tables %>%
  rename(sub = q) %>% 
  select(.imp , age, race_ethn, state, sub) %>% 
  left_join(investigation_tables %>% 
              rename(inv = q) %>% 
              select(.imp, age, race_ethn, state, inv)) %>% 
  mutate(sub_inv = sub / inv) %>% 
  filter(sub_inv<1,
         !(is.na(sub_inv)))

@@

```{r fc-inv}
fc_cond_inv_tabs<-fc_post_inv_tables %>%
  rename(fc_inv = q) %>% 
  select(.imp,age,  race_ethn, state, fc_inv) %>% 
  left_join(investigation_tables %>% 
              rename(inv = q) %>% 
              select(.imp, age, race_ethn, state, inv)) %>% 
  mutate(fc_cond_inv = fc_inv / inv)

@@

fc_cond_sub_tabs<-fc_sub_tables %>%
  rename(fc_sub = q) %>% 
  select(.imp,age,  race_ethn, state, fc_sub) %>% 
  left_join(subst_tables %>% 
              rename(sub = q) %>% 
              select(.imp, age, race_ethn, state, sub)) %>% 
  mutate(fc_cond_sub = fc_sub / sub)

@@

```{r non-icwa}
non_icwa_cond_fc<-non_icwa_tables %>%
  rename(non_icwa = q) %>% 
  select(.imp, age, race_ethn, state, non_icwa) %>% 
  left_join(fc_tables%>% 
              rename(fc = q) %>% 
              select(.imp, age, race_ethn, state, fc)) %>% 
  mutate(non_icwa_fc = non_icwa / fc)
